---
title: "ai4ph-biosignal-data-prep"
author: "Shelby Sturrock"
date: "2025-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(data.table)
setwd("/Users/dlf545/Dropbox/Teaching/USask/AI4PH/biosignal_data_ai4ph/data")
```

# 1. Read in data

*read in participant demographics*
```{r}
demo<-read.csv("participant_demog.csv")
```

*generate a list of 15 participants - approximately half male and half female with a range of ages*
```{r}
set.seed(124)
sample<-demo[sample(nrow(demo),10),] %>% dplyr::select(-race,-height_in) %>% rename(id=subj_id)
```

*Establish list of all participants with accelerometer data*
```{r}
  list<-list.files("raw_accelerometry_data", 
                   full.names=TRUE)
  list<-list[grepl(".csv",list)]
  length(list)
```

*Read in accelerometer data from each of the i number of participants (named data1, data2, ... dataX, where X is equal to i)*
Note: setting colClasses = 'character' since some a given column may be an integer for some participants but character for others, depending on whether or not they have any missing values
```{r}
  for(i in 1:length(unique(list))) { # hash and decrypted date?
    
      file<-fread(list[i],fill=TRUE,colClasses = 'character')
      file$id<-gsub(".csv","",gsub("raw_accelerometry_data/","",list[i]))
      assign(paste("data_",unique(file$id),sep=""),file)
      rm(file)
      i<-i+1
      
  }
```

# 2. Combine accelerometer data

*Combine data for all participants into one dataframe*
```{r}
  name<-"data"
  X<-mget(ls(pattern=name)) %>% bind_rows() %>% arrange(id)
  rm(list = ls()[grepl(name, ls())])
  assign(name,X)
  rm(X)
  dataOriginal<-data
```

*Remove hip accelerometer data*
```{r}
 data<-data %>% dplyr::select(id,activity,time_s,lw_x,lw_y,lw_z)
```

# 3. Merge in demographic data

only keep rows where id is in the sample file
```{r}
data<-merge(sample,data,by="id",all.x=TRUE)
```

# 4. Code `activity` as factor

```{r}
data$activity <- as.factor(data$activity)
```


```{}
data <- data %>%
	mutate(activity_cat = case_when(
		activity == "1" ~ "walking",
		activity == "2" ~ "descending stairs",
		activity == "3" ~ "ascending stairs",
		activity == "4" ~ "driving",	
		activity == "77" ~ "clapping",	
		activity == "99" ~ "non-study activity",	
		TRUE ~ "non-study activity"
	))

data$activity <- NULL
```

# 5. Write dataset

```{r}
write_csv(data, "data_clean.csv")
```

