---
title: "Data Break"
output:
      html_document:
        keep_md: true
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BiostatsUHNplus)
library(lubridate)
library(stringr)
library(seewave)
library(signal)
library(gsignal)
```

## Data Break

```{r}
data <- read_csv("data_clean.csv")
glimpse(data)

### Sort by id and time
data <- arrange(data, id, time_s)
data$id <- as.factor(data$id)

### Converting appropriate variables to factor
data <- data %>% mutate_at(c(2, 5, 6), factor)

### Creating a second level variable

data$time <- gsub("\\..*","", data$time_s)
```

## Participants

Summary statistics for the participants 

```{r}
rm_covsum_nested(data = data, id = c("id"), covs = c("gender",
"age", "weight_lbs", "age", "activity", "lw_x", "lw_y", "lw_z"))
```

We can see right away that we have strongly imblanced data for the activity. More on that later. 

## Accelerometer signals

First, we want to sort by time and id to make the plotting a bit more efficient. Unfortunately, in this sample dataset we don't get real clock time, only a index variable for the second. Second, we are going to filter out non study activity from the dataset just to make this a bit more manageable. 

__Type of activity__

* 1. Walking
* 2. Descending stairs
* 3. Ascending stairs
* 4. Driving
* 77. Clapping 
* 99. Non-study activity

```{r}
table(data$activity)

data <- dplyr::filter(data, activity != "99") ## Filter out non study activity
```

## Check frequency

```{r}
data <- data %>%
            group_by(time, id) %>%
              mutate(freq_n = n()) %>%
              ungroup()

ggplot(data, aes(x = time, y = freq_n)) + 
        geom_point() + 
        facet_wrap(~ id) + 
        theme_minimal()
```

### X Axis

```{r}
ggplot(data, aes(x = time_s, y = lw_x, colour = activity)) +
                  geom_line() + 
                  facet_wrap(~ id)

ggsave("accel_signal_x.pdf")
```

### Y Axis

```{r}
ggplot(data, aes(x = time_s, y = lw_y, colour = activity)) +
                  geom_line() + 
                  facet_wrap(~ id)


ggsave("accel_signal_y.pdf")
```

### Z Axis

```{r}
ggplot(data, aes(x = time_s, y = lw_z, colour = activity)) +
                  geom_line() + 
                  facet_wrap(~ id)

ggsave("accel_signal_z.pdf")
```

## Signal Processing 

Good resource for signal processing here: [https://cran.r-project.org/web/packages/gsignal/vignettes/gsignal.html](https://cran.r-project.org/web/packages/gsignal/vignettes/gsignal.html)

Here we have most of the data an 100Hz but some records are at a lower frequency. We will want to resample those values to create a consistent signal. It's common in the accelerometer space to resample to 30Hz. This will help later on in the analysis and is generally common practice. We can either resample the 

```{r}
data_30hz <- data %>%
              summarize(x = signal::resample(lw_x, p = 1, q = 30, d = 5),
                      y = signal::resample(lw_y, p = 1, q = 30, d = 5),
                      z = signal::resample(lw_z, p = 1, q = 30, d = 5)
                      )
```

## Features

### Time based features

Example code for a number of features available on the BEAP Lab website here [https://github.com/walkabillylab/Smartphone_accelerometers-Pocket_location/blob/master/Misc_codes/feature_generator.R
](https://github.com/walkabillylab/Smartphone_accelerometers-Pocket_location/blob/master/Misc_codes/feature_generator.R
)

One of the most common features in accelerometer analysis is the `vector magnitude` or `euclidian norm`. This feature is important because it combines information from the 3 axes. This could be important, in particular in cases where we don't know the orientation of the device. 

```{r}
data$vec_mag <- sqrt((data$lw_x^2) + (data$lw_y^2) + (data$lw_z^2))

ggplot(data, aes(x = time_s, y = vec_mag, colour = activity)) +
                  geom_line() + 
                  facet_wrap(~ id)
```

### Frequency based features

```{r}
fft_x <- fft(data$lw_x)
```

### Session info

```{r}
sessionInfo()
```

